---
title: "Preliminary Research Findings"
subtitle: "Cancel the Contract Project"
author: "By Catalyst California"
output:
  html_document:
    css: "W:\\RDA Team\\R\\cc_brandguide.css"
    toc: yes
    toc_depth: 3
    toc_float: yes
date: "`r format(Sys.time(), '%d %B, %Y')`"
---
<link href="https://fonts.googleapis.com/css2?family=Inter:wght@100;200;300;400;500;600;700;800;900&display=swap" rel="stylesheet">

```{r, include=FALSE}

knitr::opts_chunk$set(message=FALSE, echo=FALSE)

# Set up (add any additional libraries you may need)-----------------------------

library(RPostgres)
library(devtools)
library(stringr)
library(dplyr)
library(tidyr)
library(flextable)
library(ggplot2)
library(tidycensus)


#connect to postgres-------------

source("W:\\RDA Team\\R\\credentials_source.R")
con <- connect_to_db("cancel_the_contract")
con_shared <- connect_to_db("rda_shared_data")
con_fresno<- connect_to_db("eci_fresno_ripa")
con_rc<-connect_to_db("racecounts")

# data dictionary
dict<-dbGetQuery(con, "SELECT * FROM data_dictionary")

# Visualization tools (optional!)-------------------------

#add function that better fits the tables to the page

FitFlextableToPage <- function(ft, pgwidth = 6.5){

  ft_out <- ft %>% autofit()

  ft_out <- width(ft_out, width = dim(ft_out)$widths*pgwidth /(flextable_dim(ft_out)$widths))
  return(ft_out)
}


```

# I.	Executive Summary/Main Narrative

All people in the AV deserve to live in safety.  This includes freedom from police violence, as well as conditions that create well-being and opportunities to thrive.  For people of color, especially Black and Latinx youth, safety is often undermined by LA Sheriff’s Department (LASD) and high school district practices that push youth of color out of schools and into the criminal legal system (such as disproportionate suspensions, expulsions, and arrests).  More broadly, for Black and Latinx AV residents, a sense safety is undermined by inequities in other sectors, such as housing and economic insecurity. 

## A. Summary of research findings

Below is a summary of our preliminary research findings. All of the data can be explored more in-depth throughout the page: 

Education data shows that __students of color are being funneled out of the classroom through expulsions and suspensions__. Antelope Valley Union High School District (AVUHSD) suspends Black students more than to all other student groups. AVUHSD also suspends students with disabilities (SWDs) at very high rates. Conversely, these student groups also have __lower graduation rates compared to White students__. These disparities demonstrate the systemic biases that are harming students of color, and the urgent need for more investments in better school supports and counseling. 

Preliminary research findings show that __police are disproportionately stopping Black and Latinx youth AVUHSD schools__. Black students in particular are being disproportionately impacted by police relative to their overall enrollment in AVUHSD. 

The majority of people being stopped by police in AVUHSD schools are __stopped for reasons related to school fighting or possession of marijuana__. Additional stop reasons police cite include non-serious offenses such as jaywalking or violating education codes.

The results of people being stopped in AVUHSD schools further corroborates that police activity is not being used to prevent serious crime. __Most stops in AVUHSD schools result in an in field cite and release__, or some form of __referral to a parent, guardian or school administrator__. 

The racial disparities in police stops extend beyond AVUHSD schools to the greater Antelope Valley (AV) region. Police disproportionately stop people they perceive as Black relative to the general Black population in the Antelope Valley. __People perceived as Latinx by police make up nearly half of all the stops police make in the Antelope Valley__. 

# II. Background / Context of AV

## A.	Describe how the AV is a diverse place. 

```{r}

# AV Demographic tables

av_race<-dbGetQuery(con, "SELECT * FROM av_population_race")%>%
  filter(!grepl("nh_asian_wo_sa|nh_aian|nh_nhpi", race))%>%
  mutate(race_label=ifelse(race %in% 'aian', 'AIAN',
                          ifelse(race %in% 'latino', 'Latinx',
                                        ifelse(race %in% 'nh_asian', 'Asian',
                                               ifelse(race %in% 'nh_black', 'Black',
                                               ifelse(race %in% 'nh_other', 'Other',
                                                      ifelse(race %in% 'nh_twoormor', 'Multiracial',
                                                             ifelse(race %in% 'nh_white', 'White',
                                                                    ifelse(race %in% 'nhpi', 'NHPI',
                                                                           ifelse(race %in% 'sswana', 'SSWANA', "NULL"))))))))))


                       
av_age<-dbGetQuery(con, "SELECT * FROM av_population_age")

```

We define the Antelope Valley (AV) geographic region as SPA 1. The AV has a __large youth population__ overall, at __27.5% of the total population__. By comparison, 20.9% of LA county's population is age 0-17. The AV is also racially diverse, with Latinx people making up more than half of the total population. 

```{r}

df<-av_age%>%
  mutate(rate=round(rate,1))%>%
  select(age_re, total, count, rate)%>%
  rename("Age"="age_re")%>%
    rename_with(~ str_replace_all(.x, "_", " ") %>%
                str_to_lower() %>%
                str_replace("^\\w", ~ toupper(.x)))



  # Pull in county age data for comparison
  
  age_data <- dbGetQuery(con_shared, "SELECT * FROM demographics.acs_5yr_s0101_multigeo_2023 WHERE geoid = '06037' AND geolevel ='county' ")%>%
select(s0101_c01_001e	, s0101_c01_022e)%>%
    mutate(rate=s0101_c01_022e/s0101_c01_001e*100)
  
  # ##Flextable for Age Group##


  flextable(df) %>%
  set_caption('Antelope Valley Age Demographics') %>%
  theme_vanilla() %>%
  width(width=2,unit="in")
  
```

```{r}

title_text<-"The majority of residents in the AV are Latinx, White and Black"
subtitle_text<-"Race Demographics in Antelope Valley"
caption_text<-paste0("All racial groups are non-inclusive of Latinx except for AIAN, NHPI and SSWANA. AIAN=American Indian/Alaskan Native; NHPI=Native Hawaiian/Pacific Islander;SSWANA=South Asian, Southwest Asian and North Africa. \nData Source: ",dict$source[dict$indicator_short=="Race"])

#graph

ggplot(av_race, aes(x = reorder(race_label, rate), y = rate))+
  geom_bar(stat = "identity", fill = "#362178") + #to create barplot
  geom_text(aes(label = paste0(round(rate,1), "%")), #add percentage labels within the bars
          vjust=.5, #center percentage within the bar
          hjust=0,
          size = 4,
          color = "black")+ 
  scale_x_discrete(labels = ~ str_wrap(as.character(.x), 7))+
  labs(title = title_text, 
       subtitle = subtitle_text,
       caption = caption_text,
       x = "", #labels to visual
       y = "") +
  theme_bw()+
  theme(axis.text.x = element_text(angle = 0, hjust =0.5), #sets x-axis labels to be horizontal
      legend.position ="none", #removes legend
      plot.title = element_text(hjust=0.5),
      plot.subtitle = element_text(hjust=0.5),
       plot.caption = element_text(hjust = 0))+
  theme(axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=0.5, size=12), 
        strip.background = element_rect(fill = "#362178"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
panel.background = element_blank())+
  coord_flip()+
     ylim(0, max(av_race$rate) * 1.1)



```

Our education analysis for the AV focuses on the Antelope Valley Union High School District (AVUHSD). All education data for AVUHSD is for the 2024-2025 academic year. In AVUHSD there are more Black and Latinx students enrolled than White students. __Students with disabilities have a 17.8% enrollment rate__ in AVUHSD.

LA county student enrollment also has a high proportion of Latinx students at 65.2% compared to 67.2% in AVUHSD. However, the __Black student enrollment rate for LA county is only 6.7%, compared to 16.6% in AVUHSD.__ Additionally, __students with disabilities make up 14.8% of total student enrollment in LA county.__ 

```{r}

# AVUHSD Enrollment Data

enrollment <- dbGetQuery(con_shared, "SELECT * FROM education.cde_multigeo_enrollment_census_day_2024_25")

# get LAC enrollment
lac_enr<-enrollment%>%
  filter(aggregatelevel=="C" & countycode=="19" & charter == "ALL" )%>%
    filter(grepl("RE_|TA|SG_DS", reportingcategory))%>%
   mutate(total = total_enr[reportingcategory == "TA"])%>%
  filter(!grepl("TA", reportingcategory))%>%
   mutate(reportingcategory_re=ifelse(reportingcategory %in% "RE_B", "nh_black",
                                     ifelse(reportingcategory %in% "RE_I", "nh_aian", 
                                            ifelse(reportingcategory %in% "RE_A", "nh_asian", 
                                                   ifelse(reportingcategory %in% "RE_F", "nh_filipino",  
                                                          ifelse(reportingcategory %in% "RE_H", "latinx", 
                                                                 ifelse(reportingcategory %in% "RE_D", "missing_race",
                                                                        ifelse(reportingcategory %in% "RE_P", "nh_nhpi",  
                                                                               ifelse(reportingcategory %in% "RE_T", "nh_twoormor",  
                                                                                      ifelse(reportingcategory %in% "RE_W", "nh_white", 
                                                                                             reportingcategory))))))))))%>%
    mutate(race_label=ifelse(reportingcategory %in% "RE_B", "Black",
                                     ifelse(reportingcategory %in% "RE_I", "AIAN", 
                                            ifelse(reportingcategory %in% "RE_A", "Asian", 
                                                   ifelse(reportingcategory %in% "RE_F", "Filipinx",  
                                                          ifelse(reportingcategory %in% "RE_H", "Latinx", 
                                                                 ifelse(reportingcategory %in% "RE_D", "Missing race",
                                                                        ifelse(reportingcategory %in% "RE_P", "NHPI",  
                                                                               ifelse(reportingcategory %in% "RE_T", "Multiracial",  
                                                                                      ifelse(reportingcategory %in% "RE_W", "White", 
                                                                                             reportingcategory))))))))))%>%
  
  select(race_label, total_enr, total)%>%
  mutate(rate=total_enr/total*100)%>%
  arrange(-rate)


# get Antelope Valley Union High School District enrollment------------------
av_enrollment <- enrollment %>% filter(districtname == "Antelope Valley Union High" & charter == "ALL")%>%
  filter(grepl("RE_|TA", reportingcategory))%>%
  mutate(total = total_enr[reportingcategory == "TA"])%>%
  filter(!grepl("TA", reportingcategory))


# recode racial categories

av_enrollment<-av_enrollment%>%
  mutate(reportingcategory_re=ifelse(reportingcategory %in% "RE_B", "nh_black",
                                     ifelse(reportingcategory %in% "RE_I", "nh_aian", 
                                            ifelse(reportingcategory %in% "RE_A", "nh_asian", 
                                                   ifelse(reportingcategory %in% "RE_F", "nh_filipino",  
                                                          ifelse(reportingcategory %in% "RE_H", "latinx", 
                                                                 ifelse(reportingcategory %in% "RE_D", "missing_race",
                                                                        ifelse(reportingcategory %in% "RE_P", "nh_nhpi",  
                                                                               ifelse(reportingcategory %in% "RE_T", "nh_twoormor",  
                                                                                      ifelse(reportingcategory %in% "RE_W", "nh_white", 
                                                                                             reportingcategory))))))))))%>%
    mutate(race_label=ifelse(reportingcategory %in% "RE_B", "Black",
                                     ifelse(reportingcategory %in% "RE_I", "AIAN", 
                                            ifelse(reportingcategory %in% "RE_A", "Asian", 
                                                   ifelse(reportingcategory %in% "RE_F", "Filipinx",  
                                                          ifelse(reportingcategory %in% "RE_H", "Latinx", 
                                                                 ifelse(reportingcategory %in% "RE_D", "Missing race",
                                                                        ifelse(reportingcategory %in% "RE_P", "NHPI",  
                                                                               ifelse(reportingcategory %in% "RE_T", "Multiracial",  
                                                                                      ifelse(reportingcategory %in% "RE_W", "White", 
                                                                                             reportingcategory))))))))))%>%
  
  select(districtname, race_label, total_enr, total)%>%
  mutate(rate=total_enr/total*100)%>%
  arrange(-rate)

# Graph-------------------------

title_text<-"The majority of students enrolled in AVUHSD are Latinx and Black"
subtitle_text<-"Race Demographics in AVUHSD"
caption_text<-paste0("All racial groups are non-inclusive of Latinx. AIAN=American Indian/Alaskan Native; NHPI=Native Hawaiian/Pacific Islander. \nData Source: ",dict$source[dict$indicator_short=="Student Race"])

#graph

ggplot(av_enrollment, aes(x = reorder(race_label, rate), y = rate))+
  geom_bar(stat = "identity", fill = "#362178") + #to create barplot
  geom_text(aes(label = paste0(round(rate,1), "%")), #add percentage labels within the bars
          vjust=.5, #center percentage within the bar
          hjust=0,
          size = 4,
          color = "black")+ 
  scale_x_discrete(labels = ~ str_wrap(as.character(.x), 5))+
  labs(title = title_text, 
       subtitle = subtitle_text,
       caption = caption_text,
       x = "", #labels to visual
       y = "") +
  theme_bw()+
  theme(axis.text.x = element_text(angle = 0, hjust =0.5, size=5), #sets x-axis labels to be horizontal
      legend.position ="none", #removes legend
      plot.title = element_text(hjust=0.5),
      plot.subtitle = element_text(hjust=0.5),
       plot.caption = element_text(hjust = 0))+
  theme(axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=0.5, size=12), 
        strip.background = element_rect(fill = "#362178"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
panel.background = element_blank())+
  coord_flip()+
      ylim(0, max(av_enrollment$rate) * 1.1)




```

# III. AVUHSD and LA Sheriff’s Department undermine safety in the AV by harming Black and Latinx youth, and youth with disabilities through practices that push them out of school and into the criminal legal system.

## A.	Schools disproportionately push Black and Latinx youth, as well as students with disabilities out of the classroom

In AVUHSD, __only 60.4% of students with disabilities are graduating__, while 80.4% of White students are graduating. AVUHSD is graduating Latinx (79.4%) and Black students (72.7%) at __lower graduation rates__ relative to other racial groups and they are the majority of the student population. 

__Graduation rates for Latinx students in AVUHSD are lower compared to county-wide rates.__ The graduation rate for Latinx students in AVUHSD is 79.4%, while the Latinx student graduation rate in LA county is 85.6%. AVUHSD should be investing more of its resources in supporting its students academic success as opposed to removing them from classrooms. 

AVUHSD suspends 5.1% of White students compared to 6.2% of Latinx students and 18.1% of Black students. __AVUHSD is suspending Black students at more than triple the rate of White students__, showing how students of color are disproportionately losing time in the classroom. __AVUHSD also suspends 13.7 out of every 100 students with disabilities__. 

In LA county, schools also suspend Black students more than any other student groups at 5.6%, while schools county-wide suspend 1.9% of Latinx students.

AVUHSD's expulsion rates show similar trends. AVUHSD expels more Black students than any other student group, and White students are expelled the least. 


```{r}
#la county tables
susp_lac<-dbGetQuery( con_rc, "SELECT * FROM v7.arei_educ_suspension_county_2025 WHERE county_id = '06037'")%>%
   select(contains("_rate"))%>%
  pivot_longer(cols = 1:9, names_to = "var", values_to="value")

grad_lac<-dbGetQuery( con_rc, "SELECT * FROM v7.arei_educ_hs_grad_county_2025 WHERE county_id = '06037'")%>%
   select(contains("_rate"))%>%
  pivot_longer(cols = 1:9, names_to = "var", values_to="value")
  
# School analysis tables
expulsions<-dbGetQuery(con, "SELECT * FROM analysis_expulsions")
grad<-dbGetQuery(con, "SELECT * FROM analysis_graduation")
suspension<-dbGetQuery(con, "SELECT * FROM analysis_suspensions")

```

## B.	Rather than getting more time in classrooms, these students are being forced to interact more with law enforcement and funneled into the school-to-prison-pipeline 

Police stopped 904 people in AVUHSD between 2018-2023. Police responded to a call for service for 317 of the 904 people stopped, or 35% of all people stopped. In other words, police chose to stop students in 65% of stops (587 of 904).

AVUHSD is over-policing Black students. __People perceived as Black by police made up 53.2% of all stops in AVUHSD__ but their __enrollment rate in 2024-2025 is only 16.6%__. Stated another way, __for every 1,000 Black students enrolled in AVUHSD, police are stopping 131.9 Black students__. 

Police stopped 38.2 students they perceived as Latinx out of 100 people stopped. The current enrollment rate of Latinx students in AVUHSD is 67.2%. __Police are stopping people they perceive as Latinx at the second highest rate out of all student groups__.

Out of all people stopped by police in AVUHSD, people police perceived as White made up 7.9% of all stops, while the White student enrollment rate is 8.9%. 

When looking at stops that are calls for service versus stops that are not calls for service, we see that police stopped students perceived as Black more often for a stop that was not a call for service compared to a call for service. Out of all people police stopped for a stop that was not a call for service in AVUHSD, 58.2% of those people were perceived as Black. In other words, __more than half of the people police stopped without a call for service in AVUHSD are perceived as Black.__ Conversely, police stopped only 5.1% of people perceived as White for a stop that is not a call for service. 

Police stopped __23 students with disabilities__ in AVUHSD. It is important to note that police stops of students with disabilities are severely under-counted, as reported by several external sources.

The table below shows the stop reasons that police cited for the 23 people with disabilities they stopped in AVUHSD.

```{r}

stops_disability_avuhsd<-dbGetQuery(con, "SELECT * FROM analysis_stops_disability_avuhsd")
stops_race_avuhsd<-dbGetQuery(con, "SELECT * FROM analysis_stops_race_avuhsd")

```

```{r}

### EXPLORE STOP REASONS for the students with disablities ###


# incidents
lasd_incidents <- dbGetQuery(con_shared, "SELECT * FROM crime_and_justice.lasd_stops_incident_2018_2023
WHERE school_name
IN (
'Antelope Valley High',
'Antelope Valley Union High',
'Eastside High',
'Highland High',
'William J. (Pete) Knight High',
'Lancaster High',
'Littlerock High',
'Palmdale High',
'Quartz Hill High',
'R. Rex Parris High',
'Desert Winds Continuation High',
'Phoenix High Community Day',
'Phoenix Continuation'
)
ORDER BY school_name
")

# persons
lasd_persons <- dbGetQuery(con_shared, "SELECT * FROM crime_and_justice.lasd_stops_person_2018_2023")


############### Select columns of interest ########################

av_stops <- lasd_incidents %>% left_join(lasd_persons, by = "contact_id")

av_disabled <- av_stops %>% filter(no_disabilities %in% c("false", "No"))%>%
  select(reason_for_contact, reason_for_contact_narrative)%>%
  rename("Reason for contact detail"="reason_for_contact_narrative")%>%
    rename_with(~ str_replace_all(.x, "_", " ") %>%
                str_to_lower() %>%
                str_replace("^\\w", ~ toupper(.x)))

# Make table:

  flextable(av_disabled) %>%
  set_caption('Stops of students with disabilities in AVUHSD by Stop Reason') %>%
  theme_vanilla() %>%
  width(width=3.5,unit="in")


```

Examining the __stop reasons for people stopped in AVUHSD__, nearly half of stops were for __school fights__ and __smoking marijuana__, neither of which should require law enforcement.

Whether in response to a call or by officer choice,police cited a __reasonable suspicion__ that the person was engaged in criminal activity in roughly __90% of stops of students__. The other 10% of stops were because of a consensual encounter and search, to determine if a student violated a school policy, to determine whether a student is truant, an outstanding arrest warrant, a possible Ed Code violation, or traffic violation.

Whether a student __violated a school policy or Ed Code__ should be investigated by school staff as opposed to law enforcement. That was __4.2% of the stop reasons police gave for people stopped in AVUHSD.__

```{r}
stops_reason_avuhsd<-dbGetQuery(con, "SELECT * FROM analysis_stops_reason_avuhsd WHERE geography = 'Antelope Valley Union High School District'")%>%
  filter(universe=='All stops')%>%
  arrange(-stops_reason_rate)%>%
  select(geography, reason_for_contact, stops_reason_count, stops_reason_rate)%>%
  mutate(stops_reason_rate=round(stops_reason_rate,1))%>%
  rename("Count"="stops_reason_count",
         "Rate"="stops_reason_rate")%>%
      rename_with(~ str_replace_all(.x, "_", " ") %>%
                str_to_lower() %>%
                str_replace("^\\w", ~ toupper(.x)))


  flextable(stops_reason_avuhsd) %>%
  set_caption('Stops of students in AVUHSD by stop reason') %>%
  theme_vanilla() %>%
  width(width=2,unit="in")
  
 
```
 
Out of all people police stopped in AVUHSD, __13% were stopped because the officer had a reasonable suspicion that the person was engaged in criminal activity__ were due to the __student matching the description__. Stops where someone matches a description are fraught with error.

Police also stopped students for __a number unnecessary reasons__ including disrupting classwork, disruption of school activities, willfully disturbing a school zone, loud noise, litter, jaywalking, bothering other children, possessing unlawful paraphernalia. Out of all people police stopped in AVUHSD, __15.2% were stopped for fighting in a public place__ and __12.3% were stopped for possession of marijuana on school grounds__. 


```{r}

stops_reason_rs_avuhsd<-dbGetQuery(con, "SELECT * FROM analysis_stops_reason_rs_avuhsd")
 
  # change denominator to be out of all people stopped: 904 
df<-  stops_reason_rs_avuhsd%>%
  rename("Reasonable suspicion detail"="reasonable_suspicion_reason")%>%
  mutate(total=904,
         rate=round(count/904*100, 1))%>%
  arrange(-rate)%>%
  rename_with(~ str_replace_all(.x, "_", " ") %>%
                str_to_lower() %>%
                str_replace("^\\w", ~ toupper(.x)))

 flextable(df) %>%
  set_caption('Stops of people in AVUHSD for Reasonable Suspicion by Reason Detail') %>%
  theme_vanilla() %>%
  width(width=2,unit="in")

```


```{r, include=FALSE}
stops_reason_rs_codes_avuhsd<-dbGetQuery(con, "SELECT * FROM analysis_stops_reason_rs_codes_avuhsd")%>%
  slice(1:9)%>%
  select(-deactive)%>%
  mutate(total=904,
         rate=round(count/904*100, 1))

 flextable(stops_reason_rs_codes_avuhsd) %>%
  set_caption('Stops of students for Reasonable Suspicion by Offense Code') %>%
  theme_vanilla() %>%
  width(width=2,unit="in")

```

When examining what the __stop results are for people stopped in AVUHSD__, the most common stop result was an __in field cite and release__, followed by some form of __contacting a guardian or school administrator__ alone or in combination with other stop results. Police action is not necessary for student behaviors that require guardian or school administrator referrals.  

```{r}


stops_result_avuhsd<-dbGetQuery(con, "SELECT * FROM analysis_stops_result_avuhsd") %>%
  mutate(rate=round(as.numeric(rate), 1))%>%
    mutate(stop_result = str_replace(stop_result, "^\\s*\\w", ~ toupper(.x)))%>%
    rename_with(~ str_replace_all(.x, "_", " ") %>%
                str_to_lower() %>%
                str_replace("^\\w", ~ toupper(.x)))

  flextable(stops_result_avuhsd) %>%
  set_caption('Results of People Stopped in AVUHSD Schools Alone or in Combination with Other Stop Results') %>%
  theme_vanilla() %>%
  width(width=2,unit="in")
  
```

Officers can provide an offense code to accompany a stop that resulted in an in field cite and release, which gives more detail about the citation. Out of 100 people police stopped in AVUHSD, 28.4 were given an in field cite and release for __fighting or possession of marijuana__.

```{r, include=FALSE}

infield<-av_stops%>%select(contact_id, person_id, result_of_contact_in_field_cite_and_release, result_of_contact_in_field_cite_and_release_offense_codes)%>%
  group_by(result_of_contact_in_field_cite_and_release_offense_codes)%>%
  summarise(count=n())%>%
  arrange(-count)

# Grab offense codes from the Fresno RIPA db

codes<-dbGetQuery(con_fresno, "SELECT * FROM cadoj_ripa_offense_codes_2023")

codes_unique_active <- codes %>%
    # filter(is.na(offense_repealed)) %>% # remove repealed offenses
  # filter(!changes %in% c("REMOVED","REPEALED"))%>%
  group_by(offense_statute) %>%
  summarise(offense_type_of_charge=paste(offense_type_of_charge, collapse="; "),
            statute_literal_25=paste(statute_literal_25,collapse="; "))

# focus just on top counts 

infield_top<-infield%>%
  slice(1:13)%>%
  left_join(codes_unique_active, by=c("result_of_contact_in_field_cite_and_release_offense_codes"="offense_statute"))%>%
  select(-offense_type_of_charge)%>%
  rename("offense code"="result_of_contact_in_field_cite_and_release_offense_codes")

  flextable(infield_top) %>%
  set_caption('Top in field cite and release codes') %>%
  theme_vanilla() %>%
  width(width=2,unit="in")%>%
  width(j = "statute_literal_25", width =5)%>%
    width(j = "offense code", width =.5)%>%
     width(j = "count", width =2)
  

```

Instead of contributing to student safety, police are actively harming  students in AVUHSD. __Police handcuffed 19.7 people out of everyone they stopped__. Of everyone 178 students that police handcuffed in AVUHSD, 106 of those students were perceived as Black. In other words, __more than half of the people police handcuffed in AVUHSD were people police perceived as Black.__ This extreme police action is excessive and traumatizing to students, especially in a school setting. 

Police's own data show that their actions are not contributing to safety. Out of everyone police stopped in AVUHSD, 19.6% were searched, which in itself is an extreme police interaction students should not have to experience on a school campus. Out of students police searched, __53.1% of people were perceived by police as Black and 38.4% were people police perceived as Latinx.__ It is clear that police disproportionately search Black and Latinx students. Additionally, __out of everyone police searched, 72.9% yielded no contraband.__ In other words, the overwhelming majority of students police are searching on school grounds are not harboring illegal contraband. Yet the students have to live with the experience of being searched by police at school.

## C.	LASD’s racially discriminatory practices apply to AV youth outside the classroom as well

In 2023 LASD stopped 31,818 people in the AV. Of all people stopped, 6,974 people were stopped for a call for service, or only 22% of all people stopped. Conversely, __24,844 people were stopped in an officer-initiated stop__, or __78% of all people stopped__. 

Police disproportionately stop Black people in the AV relative to their total population. The data shows that  __31.6% of all people police stopped in the AV are of people police perceived as Black__, while __only 14.8% of the AV's population is Black__.

We also see __police stopping people they perceive as Latinx at high rates in the AV__, as they make up 47.4% of all people police stopped in the AV. While 53.6% of AV's population is Latinx that is still a high burden of stops being experienced by the Latinx community.

Conversely, __24% of the AV's general population is White__ yet __people perceived as White only make up 18.5% of all AV stops__, indicating that there is racial bias in who the police are stopping in the AV.

```{r}

# RIPA analysis tables: SPA 1

stops_age<-dbGetQuery(con, "SELECT * FROM analysis_stops_age")
stops_race<-dbGetQuery(con, "SELECT * FROM analysis_stops_race")
stops_age_race<-dbGetQuery(con, "SELECT * FROM analysis_stops_age_by_race")

```

Examining why police stopped people in the AV, the most common stop reason for anyone above the age of 17 is traffic violations. This is consistent with police stop behavior county-wide and in other jurisdictions. 

```{r}

stops_stops_reason_age<-dbGetQuery(con, "SELECT * FROM analysis_stops_reason_age")
stops_stops_reason_age_race<-dbGetQuery(con, "SELECT * FROM analysis_stops_reason_age_race")

```


# IV. In addition, safety for Black and Latinx AV residents is undermined by inequities in other sectors, such as economic and housing insecurity, and environment. 

## A. [PLACEHOLDER]: Short Story / Anecdote from community – can incorporate the framing of what “real safety” means to Antelope Valley community members

Antelope Valley children, like all children deserved to be cared for. There are special circumstances in the region that demand even more care. According to the UCLA Pritzker Center, the __Antelope Valley is home to the largest number of foster care placements in LA County__. The California Department of Education recognizes the importance of funding children in foster care, providing funding through its Local Control Funding Formula. Antelope Valley Union High School District should do the same and redirect investments in school police to investments in foster students and other youth deserving heightened care.

Relatedly, the Los Angeles County Department of Child and Family Services (DCFS) reports higher rates of substantiated referrals – or reports of child abuse or neglect where authorities identify reasonable cause to believe maltreatment occurred – in the Antelope Valley than in other parts of the county. At the same time, the Antelope Valley has lower rates of services supporting children and families in the child welfare system. According to the County of Los Angeles, the region has half as many domestic violence services per capita than the rest of the county.

## B. Redlining and other racist policies made by white supremacist institutions result in poorer health outcomes for BIPOC people in the AV 

Employment is an issue with many residents commuting over the hill to cities outside of the Antelope Valley for work. The type of work is unevenly distributed as __Antelope Valley Black workers are employed as an official or manager at one third the rate of White workers__. This type of salaried employment is critical for healthy and stable households.

Compared to the rest of Los Angeles County, Antelope Valley Families are more likely to earn incomes below the real cost of living. United Ways of California estimates that 41% of LA County households earn less than the real cost of living, while 51% and 52% of families in Lancaster and Palmdale, respectively, earn less than the real cost of living. In other words, __more than half of Lancaster and Palmdale households don't earn enough to cover basic costs such as food, housing, transportation, and childcare__.

The cost of housing in a once affordable area of LA County has become a problem for families. __Six in ten Antelope Valley renters now spend 30% or more of their income on housing__. This issue is particularly burdensome for AIAN families, 82% of which are rent burdened.

```{r}

# AV rent burden table
                       
av_rent_burden<-dbGetQuery(con, "SELECT subgroup, num, pop, rate FROM analysis_rent_burden")


df<-av_rent_burden%>%
  mutate(rate=round(rate,1))%>%
  rename("Burdened"="num")%>%
  rename("Percent"="rate")%>%
    rename_with(~ str_replace_all(.x, "_", " ") %>%
                str_to_lower() %>%
                str_replace("^\\w", ~ toupper(.x)))
  
  # ##Flextable rent burden##


  flextable(df) %>%
  set_caption('Antelope Valley Rent Burden') %>%
  theme_vanilla() %>%
  width(width=2,unit="in")


```

A lack of resources, especially health resources has been well documented for the Antelope Valley. __The entire Valley is a health professional shortage area for mental professionals__. Many have to wait hours for care or travel hours to cities with more accessible care. For families and children who may have mental health needs, an investment in mental health care would be vastly preferable to investments in school resource officers. Law enforcement personnel in schools lack the training to work with children with mental health issues that trained health care professionals possess.

The lack of health care resources have immediate effects on health. __The Antelope Valley has among the highest rates of infant mortality in the County__, and Black infants experience a disproportionately higher infant mortality rate compared to other racial and ethnic groups countywide.

Lack of employment, high costs, and a lack of resources have compounding effects on health, for example life expectancy. __Life expectancy in the Antelope Valley is nearly five years less than the county average__. According to the County of Los Angeles Department of Public Health the life expectancy in SPA 1 is 75.9 years, and the life expectancy in LA County as a whole is 80.4 years.

## C. The above inequities stem from both historical and present-day government decision-making . . . 

# V. Waste of public dollars on policing and school pushout versus smart investments that create true safety and wellbeing for all AV community members 
## A.	School funds are limited, and District leadership must prioritize what type of staff to hire

## B. Outside of school, in the wider community, the cities of Lancaster and Palmdale make similar decisions and prioritize even more funding towards law enforcement at the expense of true safety.

# VI. Policy solutions / recommendations:

# Conclusion

## A.	Brief recap of main arguments 

## B.	Vision for AV's future 


