---
title: "Cancel the Contract Preliminary Research Findings"
subtitle: "Draft: Internal use only"
author: "By Catalyst California"
output:
  html_document:
    css: "W:\\RDA Team\\R\\cc_brandguide.css"
    toc: yes
    toc_depth: 3
    toc_float: yes
date: "`r format(Sys.time(), '%d %B, %Y')`"
---
<link href="https://fonts.googleapis.com/css2?family=Inter:wght@100;200;300;400;500;600;700;800;900&display=swap" rel="stylesheet">

```{r, include=FALSE}

knitr::opts_chunk$set(message=FALSE, echo=FALSE)

# Set up (add any additional libraries you may need)-----------------------------

library(RPostgres)
library(devtools)
library(stringr)
library(dplyr)
library(tidyr)
library(flextable)
library(ggplot2)

#connect to postgres-------------

source("W:\\RDA Team\\R\\credentials_source.R")
con <- connect_to_db("cancel_the_contract")
con_shared <- connect_to_db("rda_shared_data")
con_fresno<- connect_to_db("eci_fresno_ripa")

# data dictionary
dict<-dbGetQuery(con, "SELECT * FROM data_dictionary")

# Visualization tools (optional!)-------------------------

#add function that better fits the tables to the page

FitFlextableToPage <- function(ft, pgwidth = 6.5){

  ft_out <- ft %>% autofit()

  ft_out <- width(ft_out, width = dim(ft_out)$widths*pgwidth /(flextable_dim(ft_out)$widths))
  return(ft_out)
}


```

# I.	Executive Summary/Main Narrative

All people in the AV deserve to live in safety.  This includes freedom from police violence, as well as conditions that create well-being and opportunities to thrive.  For people of color, especially Black and Latinx youth, safety is often undermined by LA Sheriff’s Department and High School District practices that push youth of color out of schools and into the criminal legal system (such as disproportionate suspensions, expulsions, and arrests).  More broadly, for Black and Latinx AV residents, a sense safety is undermined by inequities in other sectors, such as housing and economic insecurity.   Add qualitative here from community engagements where community members define what safety means to them 

# II. Background / Context of AV

## A.	Describe how the AV is a diverse place. 

```{r}

# AV Demographic tables

av_race<-dbGetQuery(con, "SELECT * FROM av_population_race")
av_age<-dbGetQuery(con, "SELECT * FROM av_population_age")

```

We define the Antelope Valley (AV) geographic region as SPA 1. The AV has a __large youth population__ over-all. 

```{r}

df<-av_age%>%
  mutate(rate=round(rate,1))%>%
  select(age_re, total, count, rate)

# ##Flextable for Age Group, Race##


  flextable(df) %>%
  set_caption('Antelope Valley SPA 1 Age Demographics') %>%
  theme_vanilla() %>%
  width(width=2,unit="in")
  
  
```

```{r}

title_text<-"The majority of residents in the AV are Latinx, White and Black"
subtitle_text<-"Population in Antelope Valley SPA 1"
caption_text<-paste0("Source: ",dict$source[dict$indicator_short=="Race"])

#graph

ggplot(av_race, aes(x = reorder(race, rate), y = rate))+
  geom_bar(stat = "identity", fill = "#362178") + #to create barplot
  geom_text(aes(label = paste0(round(rate,1), "%")), #add percentage labels within the bars
          vjust=.5, #center percentage within the bar
          hjust=0,
          size = 3,
          color = "black")+ 
  scale_x_discrete(labels = ~ str_wrap(as.character(.x), 5))+
  labs(title = title_text, 
       subtitle = subtitle_text,
       caption = caption_text,
       x = "", #labels to visual
       y = "") +
  theme_bw()+
  theme(axis.text.x = element_text(angle = 0, hjust =0.5), #sets x-axis labels to be horizontal
      legend.position ="none", #removes legend
      plot.title = element_text(hjust=0.5),
      plot.subtitle = element_text(hjust=0.5),
       plot.caption = element_text(hjust = 0))+
  theme(axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=0.5, size=12), #aesthetics of the barchart
        strip.background = element_rect(fill = "#362178"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
panel.background = element_blank())+
  coord_flip()


```

Our education analysis for the AV focuses on the Antelope Valley Union High School District (AVUHSD). Enrollment data for AVUHSD is for the 2024-2025 academic year. In AVUHSD there are more Black and Latinx students than White students. Students with disabilities have a 17.8% enrollment rate in AVUHSD.

```{r}

# AVUHSD Enrollment Data

enrollment <- dbGetQuery(con_shared, "SELECT * FROM education.cde_multigeo_enrollment_census_day_2024_25")

# get Antelope Valley Union High School District enrollment
av_enrollment <- enrollment %>% filter(districtname == "Antelope Valley Union High" & charter == "ALL")%>%
  filter(grepl("RE_", reportingcategory))

# recode racial categories

av_enrollment<-av_enrollment%>%
  mutate(reportingcategory_re=ifelse(reportingcategory %in% "RE_B", "nh_black",
                                     ifelse(reportingcategory %in% "RE_I", "nh_aian", 
                                            ifelse(reportingcategory %in% "RE_A", "nh_asian", 
                                                   ifelse(reportingcategory %in% "RE_F", "nh_filipino",  
                                                          ifelse(reportingcategory %in% "RE_H", "latinx", 
                                                                 ifelse(reportingcategory %in% "RE_D", "missing_race",
                                                                        ifelse(reportingcategory %in% "RE_P", "nh_nhpi",  
                                                                               ifelse(reportingcategory %in% "RE_T", "nh_twoormor",  
                                                                                      ifelse(reportingcategory %in% "RE_W", "nh_white", 
                                                                                             reportingcategory))))))))))%>%
  select(cdscode, districtname, reportingcategory_re, total_enr)%>%
  arrange(-total_enr)

  flextable(av_enrollment) %>%
  set_caption('AVUHSD Student Demographics by Race') %>%
  theme_vanilla() %>%
  width(width=2,unit="in")
  

```

# III. The AV Union High School District and LA Sheriff’s Department undermine safety in the AV by harming Black and Latinx youth, and youth with disabilities.  This occurs through practices that push them out of school and into the criminal legal system, such as suspensions, expulsions, stops and arrests. (also known as the school-to-prison pipeline).

## A.	Schools disproportionately push Black and Latinx youth, as well as SWDs out of the classroom

Students with disabilities have a __lower graduation rate at 60.4%__, demonstrating that these students should be having more time in classrooms and less time interacting with police. 
Latinx (79.4%) and Black students (72.7%) have __lower graduation rates__ relative to other racial groups and they are the majority of the student population. 

White students have a 5.1% suspension rate compared to 6.2% for Latinx students and 18.1% for Black students. 
Largest disparity in rates is for for Black students. __Black students are being suspended at more than triple the rate of White students__ this supports our narrative about students of color losing time in the classroom. 

Students with disabilities also have a high suspension rate at 13.7%. This will be important to highlight in the report especially as it relates to how are we serving our students who have different needs if we're kicking them out of classrooms. 

Expulsion rates show similar trends although rates in a vacuum are not high. Black students have the highest expulsion rate (0.5%) compared to White students (0.1%). 

```{r}

# School analysis tables
expulsions<-dbGetQuery(con, "SELECT * FROM analysis_expulsions")
grad<-dbGetQuery(con, "SELECT * FROM analysis_graduation")
suspension<-dbGetQuery(con, "SELECT * FROM analysis_suspensions")

```

## B.	Rather than getting more time in classrooms, these students are being forced to interact more with law enforcement and funneled into the school-to-prison-pipeline 


Law enforcement stopped 904 people in AVUHSD between 2018-2023. Law enforcement responded to a call for service for 317 of the 904 people stopped, or 35% of all people stopped. In other words, law enforcement chose to stop students in 65% of stops (587 of 904).

Black students are over-policed in AVUHSD. People perceived as Black made up 53.2% of all stops in AVUHSD from 2018-2023 but their enrollment rate in 2024-2025 is only 16.6%. Stated another way, for every 1,000 Black students enrolled in AVUHSD 131.9 Black students are stopped by police. 

Students perceived as Latinx made up 38.2% of all stops in AVUHSD, and the current enrollment rate of Latinx students is 67.2%. Latinx students have the second highest rate of stops out of all student groups.

Students perceived as White made up 7.9% of all stops in AVUHSD but their current enrollment rate is 8.9%. 

When looking at stops that are calls for service versus stops that are not calls for service, we see that students perceived as Black have a higher count of stops that were not a call for service compared to a call for service. Out of all people stopped for a stop that was not a call for service in AVUHSD, 58.2% of those people were perceived as Black. In other words, __more than half of the people being stopped without a call for service in AVUHSD are perceived as Black.__ Conversely, only 5.1% of people perceived as White are stopped for a stop that is not a call for service. 

Students with disabilities make up 2.5% of all stops in AVUHSD, or a total of 23 students with disabilities were stopped. The table below shows the stop reasons for the 23 people with disabilities who were stopped.


```{r}

stops_disability_avuhsd<-dbGetQuery(con, "SELECT * FROM analysis_stops_disability_avuhsd")
stops_race_avuhsd<-dbGetQuery(con, "SELECT * FROM analysis_stops_race_avuhsd")

```

```{r}

### EXPLORE STOP REASONS for the students with disablities ###


# incidents
lasd_incidents <- dbGetQuery(con_shared, "SELECT * FROM crime_and_justice.lasd_stops_incident_2018_2023
WHERE school_name
IN (
'Antelope Valley High',
'Antelope Valley Union High',
'Eastside High',
'Highland High',
'William J. (Pete) Knight High',
'Lancaster High',
'Littlerock High',
'Palmdale High',
'Quartz Hill High',
'R. Rex Parris High',
'Desert Winds Continuation High',
'Phoenix High Community Day',
'Phoenix Continuation'
)
ORDER BY school_name
")

# persons
lasd_persons <- dbGetQuery(con_shared, "SELECT * FROM crime_and_justice.lasd_stops_person_2018_2023")


############### Select columns of interest ########################

av_stops <- lasd_incidents %>% left_join(lasd_persons, by = "contact_id")

av_disabled <- av_stops %>% filter(no_disabilities %in% c("false", "No"))%>%
  select(call_for_service, reason_for_contact, reason_for_contact_narrative)

# Make table:

  flextable(av_disabled) %>%
  set_caption('Stops of students with disabilities in AVUHSD from 2018-2023') %>%
  theme_vanilla() %>%
  width(width=2,unit="in")


```

Examining the __stop reasons for people stopped in AVUHSD__, nearly half of stops were for school fights and smoking marijuana, neither of which should require law enforcement.

Whether in response to a call or by officer choice, law enforcement said there was a reasonable suspicion that the person was engaged in criminal activity in roughly 90% of stops of students. The other 10% of stops were because of a consensual encounter and search, to determine if a student violated a school policy, to determine whether a student in truant, an outstanding arrest warrant, a possible Ed Code violation, or traffic violation.

Whether a student violated a school policy or Ed Code should be investigated by school staff NOT law enforcement. That was 4.2% of the stop reasons for people stopped in AVUHSD.

```{r}
stops_reason_avuhsd<-dbGetQuery(con, "SELECT * FROM analysis_stops_reason_avuhsd WHERE geography = 'Antelope Valley Union High School District'")%>%filter(universe=='All stops')%>%arrange(-stops_reason_rate)

  flextable(stops_reason_avuhsd) %>%
  set_caption('Stops of students in AVUHSD by stop reason') %>%
  theme_vanilla() %>%
  width(width=2,unit="in")
  
 
```
 
Out of all people stopped in AVUHSD, 13% were stopped because the officer had a reasonable suspicion that the person was engaged in criminal activity were due to the student matching the a description. Stops where someone matches a description are fraught with error.

```{r}

stops_reason_rs_avuhsd<-dbGetQuery(con, "SELECT * FROM analysis_stops_reason_rs_avuhsd")
 
  # change denominator to be out of all people stopped: 904 
df<-  stops_reason_rs_avuhsd%>%
  mutate(total=904,
         rate=round(count/904*100, 1))%>%
  arrange(-rate)

 flextable(df) %>%
  set_caption('Stops of students for Reasonable Suspicion') %>%
  theme_vanilla() %>%
  width(width=2,unit="in")

```

Diving deeper into the specific reasonable suspicion stop reasons, law enforcement stopped students for a number unnecessary reasons including disrupting classwork, disruption of school activities, willfully disturbing a school zone, loud noise, litter, jaywalking, bothering other children, possessing unlawful paraphernalia. Out of all people stopped in AVUHSD, 15.2% were stopped for fighting in a public place and 12.3% were stopped for possession of marijuana on school grounds. 

```{r}
stops_reason_rs_codes_avuhsd<-dbGetQuery(con, "SELECT * FROM analysis_stops_reason_rs_codes_avuhsd")%>%
  slice(1:9)%>%
  select(-deactive)%>%
  mutate(total=904,
         rate=round(count/904*100, 1))

 flextable(stops_reason_rs_codes_avuhsd) %>%
  set_caption('Stops of students for Reasonable Suspicion by Offense Code') %>%
  theme_vanilla() %>%
  width(width=2,unit="in")

```

When examining what the __stop results are for people stopped in AVUHSD__, the most common stop result is an in field cite and release, followed by some form of contacting a guardian or school administrator. Police action is not necessary for student behaviors that require guardian or school administrator referrals.  

```{r}


stops_result_avuhsd<-dbGetQuery(con, "SELECT * FROM analysis_stops_result_avuhsd") %>%
  mutate(rate=round(as.numeric(rate), 1))

  flextable(stops_result_avuhsd) %>%
  set_caption('Results of Stops in AVUHSD Schools') %>%
  theme_vanilla() %>%
  width(width=2,unit="in")
  
```

Diving deeper into the stops that resulted in an in field cite and release, the table below shows the most common in field cite and release codes given as a result of a stop. 

```{r}

infield<-av_stops%>%select(contact_id, person_id, result_of_contact_in_field_cite_and_release, result_of_contact_in_field_cite_and_release_offense_codes)%>%
  group_by(result_of_contact_in_field_cite_and_release_offense_codes)%>%
  summarise(count=n())%>%
  arrange(-count)

# Grab offense codes from the Fresno RIPA db

codes<-dbGetQuery(con_fresno, "SELECT * FROM cadoj_ripa_offense_codes_2023")

codes_unique_active <- codes %>%
    # filter(is.na(offense_repealed)) %>% # remove repealed offenses
  # filter(!changes %in% c("REMOVED","REPEALED"))%>%
  group_by(offense_statute) %>%
  summarise(offense_type_of_charge=paste(offense_type_of_charge, collapse="; "),
            statute_literal_25=paste(statute_literal_25,collapse="; "))

# focus just on top counts 

infield_top<-infield%>%
  slice(1:13)%>%
  left_join(codes_unique_active, by=c("result_of_contact_in_field_cite_and_release_offense_codes"="offense_statute"))%>%
  select(-offense_type_of_charge)%>%
  rename("offense code"="result_of_contact_in_field_cite_and_release_offense_codes")

  flextable(infield_top) %>%
  set_caption('Top in field cite and release codes') %>%
  theme_vanilla() %>%
  width(width=2,unit="in")%>%
  width(j = "statute_literal_25", width =5)%>%
    width(j = "offense code", width =.5)%>%
     width(j = "count", width =2)
  

```

## C.	LASD’s racially discriminatory practices apply to AV youth outside the classroom as well

In 2023 LASD stopped 31,818 people in the AV. Of all people stopped, 6,974 people were stopped for a call for service, or only 22% of all people stopped. Conversely, 24,844 people were stopped in an officer-initiated stop, or 78% of all people stopped. 

When comparing the stop rates across age ranges, there are very minor differences in stop rates with a call for service and without a call for service. However, the only age range where there is a large difference between a call for service and not call for service is for those age 17 and under. Stops with call for service for youth age 17 and under is 8.2% while the stop rate for stops that were not a call for service is 1.6% for youth age 17 and under. 

Black people in the AV are disproportionately stopped relative to their total population. We see that 31.6% of all stops in the AV  are of people perceived as Black , while only 14.8% of the AV's population is Black. This disproportionality is also reflected in stops of students perceived as Black relative to Black student enrollment in AVUHSD schools. 

We also see a high rate of Latinx people being stopped in the AV, as they make up 47.4% of all stops in the AV. While 53.6% of AV's population is Latinx that is still a high burden of stops being experienced by the Latinx community.

Conversely, 24% of the AV's general population is White yet people perceived as White only make up 18.5% of all AV stops, indicating that there is racial bias in who is being stopped by police in the AV. 


```{r}

# RIPA analysis tables: SPA 1

stops_age<-dbGetQuery(con, "SELECT * FROM analysis_stops_age")
stops_race<-dbGetQuery(con, "SELECT * FROM analysis_stops_race")
stops_age_race<-dbGetQuery(con, "SELECT * FROM analysis_stops_age_by_race")

```

Examining stop reasons for people stopped in the AV, the most common stop reason for anyone above the age of 17 is traffic violations. This is consistent with police stop behavior in LA county and other jurisdictions. 

```{r}

stops_stops_reason_age<-dbGetQuery(con, "SELECT * FROM analysis_stops_reason_age")
stops_stops_reason_age_race<-dbGetQuery(con, "SELECT * FROM analysis_stops_reason_age_race")

```
