---
title: ""
output:
  html_document:
    includes:
      in_header: "../header.html"
    mathjax: null
---

```{r, include=FALSE}
# Check if document is being run as a child, if yes don't apply styling again
is_child <- !is.null(knitr::opts_knit$get("parent.frame"))
if (is_child) {
  knitr::opts_chunk$set(include = FALSE)
}
```

```{r, include=FALSE}

# Set up (add any additional libraries you may need)-----------------------------

library(RPostgres)
library(devtools)
library(stringr)
library(dplyr)
library(tidyr)
library(flextable)
library(ggplot2)

#connect to postgres-----------------

source("W:\\RDA Team\\R\\credentials_source.R")
con <- connect_to_db("cancel_the_contract")
con_shared <- connect_to_db("rda_shared_data")

# Visualization tools (optional!)-------------------------

#add function that better fits the tables to the page

FitFlextableToPage <- function(ft, pgwidth = 6.5){

  ft_out <- ft %>% autofit()

  ft_out <- width(ft_out, width = dim(ft_out)$widths*pgwidth /(flextable_dim(ft_out)$widths))
  return(ft_out)
}

# See "W:\Project\RJS\CTC\Github\JZ\cancel_the_contract\RDA Sense Making\sensemaking_jz.Rmd" for one ggplot and one flextable example
# if you want to visualize tables quickly


# Guiding questions-------------------------

#1) What findings are you seeing that are interesting or surprising to you?
#2) What findings are you seeing that supplement OR contradict our narrative? 
#3) What additional questions do you have?
#4) What visualization ideas (chart types, figures to highlight, etc.) do you have? 

```

## CR General notes

Nearly half of stops were for school fights and smoking marijuana, neither of which should require law enforcement.

Whether in response to a call or by officer choice, law enforcement said there was a reasonable suspicion that the person was engaged in criminal activity in roughly 90% of stops of students. The other 10% of stops were because of a consensual encounter and search, to determine if a student violated a school policy, to determine whether a student in truant, an outstanding arrest warrant, a possible Ed Code violation, or traffic violation.

Whether a student violated a school policy or Ed Code should be investigated by school staff NOT law enforcement. That was 4.2% of the stop reasons for people stopped in AVUHSD.

```{r}
stops_reason_avuhsd<-dbGetQuery(con, "SELECT * FROM analysis_stops_reason_avuhsd WHERE geography = 'Antelope Valley Union High School District'")%>%filter(universe=='All stops')%>%arrange(-stops_reason_rate)

  flextable(stops_reason_avuhsd) %>%
  set_caption('Stops of students in AVUHSD by stop reason') %>%
  theme_vanilla() %>%
  width(width=2,unit="in")
  
 
```
 

15% of stops where the officer had a reasonable suspicion that the person was engaged in criminal activity were due to the student matching the a description. Stops where someone matches a description are fraught with error.

```{r}
stops_reason_rs_avuhsd<-dbGetQuery(con, "SELECT * FROM analysis_stops_reason_rs_avuhsd")%>%
  mutate(rate=round(rate*100, 1))%>%
  arrange(-rate)

 flextable(stops_reason_rs_avuhsd) %>%
  set_caption('Stops of students for Reasonable Suspicion') %>%
  theme_vanilla() %>%
  width(width=2,unit="in")

```

Diving deeper into the specific reasonable suspicion stop reasons, law enforcement stopped students for a number unnecessary reasons including disrupting classwork, disruption of school activities, willfully disturbing a school zone, loud noise, litter, jaywalking, bothering other children, possessing unlawful paraphernalia.

However, when  you look at the most  common reasonable suspicion reason offense codes, we can see that most stops of people for reasonable suspicion in AVUHSD were for posession of marijuana, fighting or some form of battery. 

```{r}
stops_reason_rs_codes_avuhsd<-dbGetQuery(con, "SELECT * FROM analysis_stops_reason_rs_codes_avuhsd")%>%
  slice(1:9)%>%
  select(-deactive)

 flextable(stops_reason_rs_codes_avuhsd) %>%
  set_caption('Stops of students for Reasonable Suspicion by Offense Code') %>%
  theme_vanilla() %>%
  width(width=2,unit="in")

```

Very different not school with traffic violation being number one.

```{r}


# Supplemental reason analysis using word count / narrative column method. But I think for more precision we should primarily rely on the tables above for the stop reason analysis.

stops_reason_reason_narrative_avuhsd<-dbGetQuery(con, "SELECT * FROM analysis_stops_reason_and_narrative_avuhsd")
stops_reason_narrative_avuhsd<-dbGetQuery(con, "SELECT * FROM analysis_stops_narrative_avuhsd")


```


Outside of AVUHSD and looking at the reason why people are stopped in the AV, the most common stop reason for anyone above the age of 17 is traffic violations. This is consistent with police stop behavior in LA county and other jurisdictions. 

```{r}

stops_stops_reason_age<-dbGetQuery(con, "SELECT * FROM analysis_stops_reason_age")
stops_stops_reason_age_race<-dbGetQuery(con, "SELECT * FROM analysis_stops_reason_age_race")

```


