---
title: ""
output:
  html_document:
    includes:
      in_header: "../header.html"
    mathjax: null
---

```{r, include=FALSE}
# Check if document is being run as a child, if yes don't apply styling again
is_child <- !is.null(knitr::opts_knit$get("parent.frame"))
if (is_child) {
  knitr::opts_chunk$set(include = FALSE)
}
```

```{r, include=FALSE}

# Set up (add any additional libraries you may need)-----------------------------

library(RPostgres)
library(devtools)
library(stringr)
library(dplyr)
library(tidyr)
library(flextable)
library(ggplot2)


# Visualization tools (optional!)-------------------------

#add function that better fits the tables to the page

FitFlextableToPage <- function(ft, pgwidth = 6.5){

  ft_out <- ft %>% autofit()

  ft_out <- width(ft_out, width = dim(ft_out)$widths*pgwidth /(flextable_dim(ft_out)$widths))
  return(ft_out)
}

# See "W:\Project\RJS\CTC\Github\JZ\cancel_the_contract\RDA Sense Making\sensemaking_jz.Rmd" for one ggplot and one flextable example
# if you want to visualize tables quickly

# See this script for examples of how to use the visual functions:
## W:\\Project\\RJS\CTC\\Github\\JZ\\cancel_the_contract\\Visuals\\visual_fx.R

# Prompting questions-------------------------

#1) What findings are you seeing that are interesting or surprising to you?
#2) What findings are you seeing that supplement OR contradict our narrative? 
#3) Is there a particular denominator or cut within the analysis that is more compelling?
#4) What additional questions do you have?
#5) What visualization ideas (chart types, figures to highlight, etc.) do you have? 

# Narrative outline: 
# https://catalystcalifornia.sharepoint.com/:w:/s/Portal/EXd-ZVrmsfdKtVkYfZJlqOoBhMIkC9oLjgHHDoKvoct4aQ?e=ly1qcg



```

## JZ General notes

* General notes and observations outside of specific analyses

```{r}

#connect to postgres
source("W:\\RDA Team\\R\\credentials_source.R")
con <- connect_to_db("cancel_the_contract")
con_shared <- connect_to_db("rda_shared_data")

# data dictionary
dict<-dbGetQuery(con, "SELECT * FROM data_dictionary")


# School analysis tables
expulsions<-dbGetQuery(con, "SELECT * FROM analysis_expulsions") #JZ
grad<-dbGetQuery(con, "SELECT * FROM analysis_graduation") # JZ
suspension<-dbGetQuery(con, "SELECT * FROM analysis_suspensions") # JZ


# RIPA analysis tables: SPA 1

stops_age<-dbGetQuery(con, "SELECT * FROM analysis_stops_age") # AV
stops_race<-dbGetQuery(con, "SELECT * FROM analysis_stops_race") # AV
stops_age_race<-dbGetQuery(con, "SELECT * FROM analysis_stops_age_by_race") # AV

stops_stops_reason_age<-dbGetQuery(con, "SELECT * FROM analysis_stops_reason_age") # CR
stops_stops_reason_age_race<-dbGetQuery(con, "SELECT * FROM analysis_stops_reason_age_race") # CR


# RIPA analysis tables: AVUHSD
stops_disability_avuhsd<-dbGetQuery(con, "SELECT * FROM analysis_stops_disability_avuhsd") # AV
stops_race_avuhsd<-dbGetQuery(con, "SELECT * FROM analysis_stops_race_avuhsd") # AV

stops_reason_avuhsd<-dbGetQuery(con, "SELECT * FROM analysis_stops_reason_avuhsd")  # CR
stops_reason_rs_avuhsd<-dbGetQuery(con, "SELECT * FROM analysis_stops_reason_rs_avuhsd") # CR
stops_reason_rs_codes_avuhsd<-dbGetQuery(con, "SELECT * FROM analysis_stops_reason_rs_codes_avuhsd") # CR

stops_result_avuhsd<-dbGetQuery(con, "SELECT * FROM analysis_stops_result_avuhsd") # JZ


stops_reason_reason_narrative_avuhsd<-dbGetQuery(con, "SELECT * FROM analysis_stops_reason_and_narrative_avuhsd") # CR
stops_reason_narrative_avuhsd<-dbGetQuery(con, "SELECT * FROM analysis_stops_narrative_avuhsd") # CR


```

## Antelope Valley general demographics

* Notes
* Ideas about visualizations

```{r}

# AV Demographic tables

av_race<-dbGetQuery(con, "SELECT * FROM av_population_race")
av_age<-dbGetQuery(con, "SELECT * FROM av_population_age")


```

### Antelope Valley population by race

* The top 3 racial groups in the AV are Latinx, White and then Black. 
* It would seem like the majority of Asians in the AV are not South Asian (comparing the NH Asian rate vs NH Asian WITHOUT South Asian rate) so there probably is not a very large South Asian population in the AV
* I think we will probably want a basic bar graph of the general race population in the report for context as it relates to other indicators even non-RIPA indicators.

```{r}

title_text<-"The majority of residents in the AV are Latinx, White and Black"
subtitle_text<-"Population in Antelope Valley SPA 1"
caption_text<-paste0("Source: ",dict$source[dict$indicator_short=="Race"])

#graph

ggplot(av_race, aes(x = reorder(race, rate), y = rate))+
  geom_bar(stat = "identity", fill = "#362178") + #to create barplot
  geom_text(aes(label = paste0(round(rate,1), "%")), #add percentage labels within the bars
          vjust=.5, #center percentage within the bar
          hjust=0,
          size = 3,
          color = "black")+ 
  scale_x_discrete(labels = ~ str_wrap(as.character(.x), 5))+
  labs(title = title_text, 
       subtitle = subtitle_text,
       caption = caption_text,
       x = "", #labels to visual
       y = "") +
  theme_bw()+
  theme(axis.text.x = element_text(angle = 0, hjust =0.5), #sets x-axis labels to be horizontal
      legend.position ="none", #removes legend
      plot.title = element_text(hjust=0.5),
      plot.subtitle = element_text(hjust=0.5),
       plot.caption = element_text(hjust = 0))+
  theme(axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=0.5, size=12), #aesthetics of the barchart
        strip.background = element_rect(fill = "#362178"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
panel.background = element_blank())+
  coord_flip()


```

### Antelope Valley population by age

* The AV has a pretty large youth population that will be worth highlighting in the report. 
* We could probably even get some more AV youth context for the report from the Region 5 State of the Child report. 
* I am not sure if this has to be a report visual it could just be context in the written part of the report.

```{r}

df<-av_age%>%
  mutate(rate=round(rate,1))%>%
  select(age_re, total, count, rate)

# ##Flextable for Age Group, Race##


  flextable(df) %>%
  set_caption('Antelope Valley SPA 1 Age Demographics') %>%
  theme_vanilla() %>%
  width(width=2,unit="in")
  
  
```

## AVUHSD general demographics

* Interesting to see that there are more Black students than White. So if we have RIPA AVUHSD indicators with enrollment denominators something to consider and what arguments of disproportionately we can make. 
* Also given high Latinx/Black student enrollment need to be mindful of how we frame AVUHSD stops of these groups. 

```{r}

# AVUHSD Enrollment Data

enrollment <- dbGetQuery(con_shared, "SELECT * FROM education.cde_multigeo_enrollment_census_day_2024_25")

# get Antelope Valley Union High School District enrollment
av_enrollment <- enrollment %>% filter(districtname == "Antelope Valley Union High" & charter == "ALL")%>%
  filter(grepl("RE_", reportingcategory))

# recode racial categories

av_enrollment<-av_enrollment%>%
  mutate(reportingcategory_re=ifelse(reportingcategory %in% "RE_B", "nh_black",
                                     ifelse(reportingcategory %in% "RE_I", "nh_aian", 
                                            ifelse(reportingcategory %in% "RE_A", "nh_asian", 
                                                   ifelse(reportingcategory %in% "RE_F", "nh_filipino",  
                                                          ifelse(reportingcategory %in% "RE_H", "latinx", 
                                                                 ifelse(reportingcategory %in% "RE_D", "missing_race",
                                                                        ifelse(reportingcategory %in% "RE_P", "nh_nhpi",  
                                                                               ifelse(reportingcategory %in% "RE_T", "nh_twoormor",  
                                                                                      ifelse(reportingcategory %in% "RE_W", "nh_white", 
                                                                                             reportingcategory))))))))))%>%
  select(cdscode, districtname, reportingcategory_re, total_enr)

```

```{r}

# School analysis tables
expulsions<-dbGetQuery(con, "SELECT * FROM analysis_expulsions")
grad<-dbGetQuery(con, "SELECT * FROM analysis_graduation")
suspension<-dbGetQuery(con, "SELECT * FROM analysis_suspensions")

```

## Stop results in AVUHSD

* Chauncee looked at an exported list of 'in field cite and release' result codes but wondering if we need to add another analysis table with actual codes and rates of the different codes. 

```{r}
stops_result_avuhsd<-dbGetQuery(con, "SELECT * FROM analysis_stops_result_avuhsd") # JZ


```